
# Supplementary material
## Williams et al. 2015 Human, oceanographic and habitat drivers of Central and Western Pacific Coral Reef Fish Assemblages

========================================================

This is the R code used to run the analysis in the paper. 
```{r, message=FALSE}
rm(list=ls())   # clean workspace
library(MuMIn)  # needed for model averaging
library(mgcv)   # needed for GAM
setwd("~/Downloads/") # set working directory
load(file="WilliamsetalPLosSONE2015.rdata") # reads in dataset z
```

### Explanation of dataframe variables
* REGION, ISLAND .. region and island names (e.g. "MHI" and"Oahu")  
* PRIM  Biomass (g/m2) of Primary Consumers  
* SECO  Biomass (g/m2) of Secondary Consumers  
* PLNK  Biomass (g/m2) of Planktivores  
* PISC_noSJ  Biomass (g/m2) of piscivores, excluding sharks and jacks    
* ALLF_noSJ  Biomass (g/m2) of all fishes, excluding sharks and jacks  
* HC      Visually estimated Hard Coral Cover (%)  
* CX		 Visually estiamted mean substrate height within survey cylinders (m)  
* CHL		Oceanic productivity, i.e. long-term mean of satelitte derived Chl-a of waters surrounding the reef areas  
* SSTL		Climatological low sea surface temperature, satelitte derived   
* HUM		Square-root transformed human population density per reef   
                area (ie humans rsident ont he island divided by area of         forereef)  
* HDIST		Square-root transformed 'distant' human population per reef area (i.e. humans living within 200km of the site but not resident on the island, divided by the area of forereef)  
* AT		Atoll? (True/False)
* WV  	        Long-term mean estimated island-scale wave energy 

### MuMIn routines to generate and rank all possible models, and to model average the selected models

```{r, message=FALSE}
MXPARAMS<-7  # maximum # of parameters in the model 
BASE_K<-5    # limit to number of knots

#####   Example code for ALLF_noSJ
z$BIO<-z$ALLF_noSJ

#define full model
M1<-gam(BIO ~ s(CX, k=BASE_K) 
        + s(CHL, k=BASE_K) 
        + s(HC, k=BASE_K) 
        + s(HUM, k=BASE_K) 
        + s(HDIST, k=BASE_K) 
        + s(SSTL, k=BASE_K) 
        + s(WV, k=BASE_K) 
        + AT, 
        data=z, family=Gamma(link=log))

# run all possible combinations, but do not include both SSTL and WV in the same model
options(na.action="na.fail") # prevent fitting models to different datasets

M.set<-dredge(M1,beta=FALSE, rank="AICc", subset=!(`s(SSTL, k = BASE_K)`  &&  `s(WV, k = BASE_K)`), m.max=MXPARAMS, extra = alist(AIC, "R^2", "adjR^2"))


#head(M.set,12) #view top models, and variable importance
#round(importance(M.set),3)       

top.models <- get.models(M.set, weight > .05) #select all models with weight > 0.05

# use model averaging of all selected top models
M.avg<-model.avg(top.models) # get averaged coefficients
```

### Code to generate predictions in absence of humans

A predictor data set is created from the full data set, and set the two human population density variables to 0
```{r, message=FALSE}

PD<-z # duplicate dataset
PD$HUM<-0 # set local human population to 0
PD$HDIST<-0 # set distance human population to 0

# run predict function on the averaged mdoel (generated above)
PRED<-predict(M.avg, newdata=PD, se.fit=TRUE, backtransform=FALSE, type="response")

#generate prediction output data frame
PO<-z[,c("REGION", "ISLAND", "PRIM", "SECO", "PLNK", "PISC_noSJ", "ALLF_noSJ")]
PO$REGION<-as.character(PO$REGION)
PO$fit<-PRED$fit
PO$se<-PRED$se.fit
```
